----
title: "Project Proposal"
author: "Echo Chen, Andrew Kroening, Pooja Kabber, Dingkun Yang"
output: pdf_document
date: "November 22nd, 2022"
---

# Report Draft
## 1.load data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
packages <- (c("table1", "ggplot2", "xtable", "pander","ggcorrplot", "tidyverse", "stargazer", "ggfortify", "caTools", "car", "corrplot", "gridExtra"))
# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

```{r read_data, echo=FALSE, include=FALSE}
df_life_expectancy <- read.csv("/Users/andrewkroening/Desktop/orange-modeling-project/Datasets/Life_Expectancy_Data.csv")
df_life_expectancy <- data.frame(df_life_expectancy)
df_life_expectancy$Status <- factor(df_life_expectancy$Status)
df_life_expectancy_2014 <- df_life_expectancy[df_life_expectancy$Year == 2014, ]
```
## 2. clean data 
```{r}
# high VIFS
mod.linear <- lm(df_Life.expectancy~ ., data = subset(data, select =-c(Status)))
vifs <- data.frame(vif(mod.linear))

set_plot_dimensions(16,8)
ggplot(vifs, aes(y=vif.mod.linear., x=row.names(vifs))) + 
    geom_bar(aes(fill=vif.mod.linear.>5),stat="identity")+
    scale_y_continuous(trans = "sqrt",  breaks = c(5, 10, 50, 100))+
    geom_hline(yintercept = 5, colour = "red") + 
    ggtitle("VIF per feature") +
    xlab("Featurs") + ylab("VIF") +
    theme(axis.text.x=element_text(angle=20, hjust=1))+
    theme(text = element_text(size = 18))+
    scale_fill_brewer(palette="Dark2")
```
```{r}
# drop high VIFS
data_EDA <- subset(df_Life.expectancy, select = -c(infant.deaths))
data_EDA <- subset(data_EDA, select = -c(GDP))
data_EDA <- subset(data_EDA, select = -c(thinness..1.19.years))
```
```{r}
# check correlation matrix and VIF again
set_plot_dimensions(16,10)
corr <- round(cor(subset(data_EDA, select =-c(Status))), 3)
ggcorrplot(corr,type = "upper", lab = TRUE, outline.color = "black", lab_size = 4, legend.title = "Correlation")
mod.linear <- lm(Life.expectancy~ ., data = subset(data_EDA, select =-c(Status)))
vifs <- data.frame(vif(mod.linear))

set_plot_dimensions(16,8)
ggplot(vifs, aes(y=vif.mod.linear., x=row.names(vifs))) + 
    geom_bar(aes(fill=vif.mod.linear.<5),stat="identity")+
    scale_y_continuous(trans = "sqrt",  breaks = c(5, 10, 50, 100))+
    geom_hline(yintercept = 5, colour = "red") + 
    ggtitle("VIF per feature") +
    xlab("Featurs") + ylab("VIF") +
    theme(axis.text.x=element_text(angle=20, hjust=1))+
    theme(text = element_text(size = 18))+
    scale_fill_brewer(palette="Dark2")
```


## 3. choose variables( R & AIC )
```{r}
# plot 
regfit.best <- regsubsets(Life.expectancy~., data= data_EDA, nvmax = 16)
reg.summary <- summary(regfit.best)
# adjusted-R^2 with its largest value
plot(reg.summary$adjr2,xlab="Number of Variables",ylab="Adjusted Rsq",type="l")
which.max(reg.summary$adjr2)
points(15,reg.summary$adjr2[15], col="red",cex=2,pch=20)

# AIC with its smallest value
plot(reg.summary$aic,xlab="Number of Variables",ylab="AIC",type='l')
which.min(reg.summary$aic)
points(12,reg.summary$aic[12],col="red",cex=2,pch=20)
# forward backward?
```

## 4. transformation (missing & skewed)

## 5. select model (check asumptions)
```{r}
data_FS <- subset(data_EDA, select=c(Life.expectancy,Status, Adult.Mortality, percentage.expenditure,
                                     Hepatitis.B, Polio, BMI, thinness.5.9.years, Measles,
                                     Diphtheria,HIV.AIDS,Income.composition.of.resources,Schooling))
set.seed(123)

sample <- sample(c(TRUE, FALSE), nrow(data_FULL), replace=TRUE, prob=c(0.70,0.30))
train <- data_FULL[sample, ]
x.test <-data_FULL[!sample, ]
y.test <- data_FULL[!sample, ]$Life.expectancy
model.full<- lm(Life.expectancy~., data = train)
summary(model.full)
```
```{r}
pred <- predict(model.full, newdata=x.test)
rmse(pred,y.test)
summary(model.full)$adj.r.squared
par(mfrow=c(2,2))
plot(model.full)
```
The linearity assumptions are met for the full model We can notice that all linear regression assumptions are roughly met:

- The Residuals vs Fitted plot

There is no pattern in the residual plot. we can assume linear relationship between the predictors and the outcome variables. (linearity assumption)

- The Scale-Location plot

The residuals are spread equally across the ranges of predictors. (homoscedasticity assumption)

The Normal Q-Q plot
All the points spread along the reference line equally, it’s good enough to assume that the data has the normality of the residuals.(normality of the residuals assumption)

The Residuals vs Leverage
Any point in the 0.5 cook distance boundary are influential. However, there are no influential points here. (linearity assumption)

## 6.models with conclusions

```{r}
set.seed(123)

sample <- sample(c(TRUE, FALSE), nrow(data_FS), replace=TRUE, prob=c(0.70,0.30))
train <- data_FS[sample, ]
x.test <-data_FS[!sample, ]
y.test <- data_FS[!sample, ]$Life.expectancy
model.FS <- lm(Life.expectancy~., data = train)
summary(model.FS)
#prediction ?
pred <- predict(model.EDA, newdata=x.test)
rmse(pred,y.test)
summary(model.EDA)$adj.r.squared
par(mfrow=c(2,2))
plot(model.EDA)
# cheack assumption
pred <- predict(model.FS, newdata=x.test)
rmse(pred,y.test)
summary(model.FS)$adj.r.squared
par(mfrow=c(2,2))
plot(model.FS)
```

