---
title: "Primary relationship of interest"
author: "Team Orange"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
packages <- (c("tidyverse", "stargazer", "ggfortify", "caTools", "car", "corrplot", "gridExtra"))

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

## Primary relationship of interest


```{r, echo=FALSE, results="asis", header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
# Import the data and look at the first six rows
le <- read.csv(file = 'Life Expectancy Data.csv')
#head(le)
le14 <- le[which(le$Year==2014),]
#str(le14)
#names(le14)
#head(le14)
#summary(le14)
# fit <- lm(Life.expectancy ~ Adult.Mortality+infant.deaths+under.five.deaths+percentage.expenditure+Hepatitis.B+Measles+BMI+Polio+HIV.AIDS+Diphtheria+GDP+Schooling+ Income.composition.of.resources+Status+Population+thinness..1.19.years+thinness.5.9.years, data=le14)
# stargazer(fit,header=FALSE, type="latex", 
#           no.space = TRUE,
#           report = ('vcsp*'), single.row = TRUE,
#           column.sep.width = "0.2pt",
#           font.size = "small",
#           title="Regression Summary")
```


```{r, echo=FALSE, results="asis", header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
le <- read.csv(file = 'Life Expectancy Data.csv')
le14 <- le[which(le$Year==2014),]
#head(le)
```

Full correlation map (Maybe in appendix):

```{r cor, fig.width=8, fig.height=8, fig.show='hold', out.width="90%", fig.align='center', results="asis", echo=FALSE, header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
le14_cont <- le14[,c("Life.expectancy", "Adult.Mortality", "infant.deaths","Alcohol","percentage.expenditure", "Hepatitis.B","Measles","BMI","under.five.deaths","Polio","Total.expenditure","Diphtheria" ,"HIV.AIDS","GDP","Population","thinness..1.19.years","thinness.5.9.years","Income.composition.of.resources",
"Schooling")]
le14_cont_clean <- na.omit(le14_cont)
le14_cont_clean.cor <- cor(le14_cont_clean)
par(mfrow = c(1, 1))
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(le14_cont_clean.cor, method="color", col=col(200),  
     diag=FALSE, # tl.pos="d", 
         type="upper", order="hclust", 
     # title=title, 
         addCoef.col = "black", # Add coefficient of correlation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.05, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
     tl.cex=0.7,
     number.cex=0.7,
     mar=c(0,0,0,0)
         )
```

### After piror selection:

When building a model, it is not wise use all variables when they are highly correlated with each other. To represent immunization coverage, among "Hepatitis.B","Polio", "Diphtheria tetanus toxoid and pertussis (DTP3)", we decide to use "polio", since it has the highest correlation between Life Expectancy. Similarly, we see extremely high correlation between GDP and percentage expenditure, but both of them would have meaningful interpretation, we may want to decide which one goes to our final model when we conduct model selection.

On the one hand, with domain knowledge, we know "Adult.Mortality", "infant.deaths" and "under.five.deaths" variables are directly correlated to Life Expectancy, we choose to drop them from the predictor variable list; on the other hand, we are interested in "HIV.AIDS" variable (Deaths per 1 000 live births HIV/AIDS (0-4 years), .... (please add some reasons here)

We would like to omit the variables have Low correlation between Life Expectancy: "Measles"; however, we do want to include "population" because of our interest. 

As for categorical variable, we would like to keep country status (developing/developed) as one of the predictors


```{r cor1, fig.width=6, fig.height=5, fig.show='hold', out.width="80%", fig.align='center', results="asis", echo=FALSE, header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}

le14_piror <- le14_cont_clean[,c("BMI","Life.expectancy", "percentage.expenditure", "Polio","Total.expenditure" ,"HIV.AIDS","GDP","Population","Income.composition.of.resources",
"Schooling")]

le14_cont_piror.cor <- cor(le14_piror)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(le14_cont_piror.cor, method="color", col=col(200),  
     diag=FALSE, # tl.pos="d", 
         type="upper", order="hclust", 
     # title=title, 
         addCoef.col = "black", # Add coefficient of correlation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.05, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
     tl.cex=0.75,
     number.cex=0.75,
     mar=c(0,0,0,0)
         )

```

**Potential predictor variables:**

- *Percentage Expenditure* or *GDP*
- *BMI*
- *Polio*
- *HIV/AIDS*
- *Total.expenditure*
- *Schooling*
- *Income Composition of Resources*
- *Country Status (Developed/Developing)*
- *Population*



\newpage

### Transformation if needed when modeling

When checking box plots of all variables, we find 4 variables may need some transformation.

```{r, fig.width=7, fig.height=5, fig.show='hold', out.width="85%", fig.align='center', results="asis", echo=FALSE, header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
## If interested, Comment out to see whats going on ##
# c <- ggplot(le14, aes(x=Life.expectancy))+
#   geom_boxplot(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "Life Expectancy in age")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))
# c1 <- ggplot(le14, aes(x=Polio))+
#   geom_boxplot(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "Polio (Pol3) immunization coverage among 1-year-olds (%)")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))
# c2 <- ggplot(le14, aes(x=BMI))+
#   geom_boxplot(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "Average Body Mass Index of entire population")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))
c3 <- ggplot(le14, aes(x=percentage.expenditure))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "Expenditure on health as a percentage of Gross Domestic Product per capita(%)")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))
# c4 <- ggplot(le14, aes(x=Total.expenditure))+
#   geom_boxplot(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "General government expenditure on health as a percentage of total government expenditure (%)")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))
c5 <- ggplot(le14, aes(x=HIV.AIDS))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "Deaths per 1 000 live births HIV/AIDS (0-4 years)")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))
c6 <- ggplot(le14, aes(x=GDP))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "Gross Domestic Product per capita (in USD)")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))
c7 <- ggplot(le14, aes(x=Population))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "Population of the country")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))
# c8 <- ggplot(le14, aes(x=Income.composition.of.resources))+
#   geom_boxplot(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "Human Development Index in terms of income composition of resources (index ranging from 0 to 1)")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))
# c9 <- ggplot(le14, aes(x=Schooling))+
#   geom_boxplot(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "Number of years of Schooling(years)")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))

### All Box Plot, if comment out ###
# grid.arrange(c,c1,c2,c3,c4,c5,c6,c7,c8,c9)

grid.arrange(c3,c5,c6,c7)
```

We may want to use log transformation for population and GDP, since the magnitude of gaps are huge. But for other two, we need more investigation due to the difficulty of interpretation.


The box plots and histograms of population and GDP after log transformation:

```{r, fig.width=7, fig.height=4, fig.show='hold', out.width="85%", fig.align='center', results="asis", echo=FALSE, header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}

# c3 <- ggplot(le14, aes(x=percentage.expenditure))+
#   geom_histogram(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "Expenditure on health as a percentage of Gross Domestic Product per capita(%)")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))
# c5 <- ggplot(le14, aes(x=HIV.AIDS))+
#   geom_histogram(color="darkblue", fill="lightblue")+
#   labs(#title="Box Plot of Polio for Children Under 3 Years Old",
#       x = "Deaths per 1 000 live births HIV/AIDS (0-4 years)")+
#   theme(plot.title = element_text(size=10),
#         axis.title = element_text(size=6))
c6 <- ggplot(le14, aes(x=log10(GDP)))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "ln(Gross Domestic Product per capita (in USD))")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))
c7 <- ggplot(le14, aes(x=log10(Population)))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "ln(Population of the country)")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))

c6_1 <- ggplot(le14, aes(x=log10(GDP)))+
  geom_histogram(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "ln(Gross Domestic Product per capita (in USD))")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))
c7_1 <- ggplot(le14, aes(x=log10(Population)))+
  geom_histogram(color="darkblue", fill="lightblue")+
  labs(#title="Box Plot of Polio for Children Under 3 Years Old",
      x = "ln(Population of the country)")+
  theme(plot.title = element_text(size=10),
        axis.title = element_text(size=6))

grid.arrange(c6,c7,c6_1,c7_1,ncol=2)
```


\newpage

 (For our reference)

After selection Model:

```{r, echo=FALSE, results="asis", header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
fit2 <- lm(Life.expectancy ~ BMI+log10(GDP)+percentage.expenditure+Polio+HIV.AIDS+Total.expenditure+log10(Population)+Income.composition.of.resources+Status+Schooling, data=le14)
stargazer(fit2,header=FALSE, type="latex", 
          no.space = TRUE,
          report = ('vcsp*'), single.row = TRUE,
          column.sep.width = "0.2pt",
          font.size = "small",
          title="Regression Summary")
par(mfrow = c(2, 2))
plot(fit2)
```

Potential Final Model :

```{r, echo=FALSE, results="asis", header=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
fit2 <- lm(Life.expectancy ~ log10(GDP)+HIV.AIDS+Income.composition.of.resources+Status, data=le14)
stargazer(fit2,header=FALSE, type="latex", 
          no.space = TRUE,
          report = ('vcsp*'), single.row = TRUE,
          column.sep.width = "0.2pt",
          font.size = "small",
          title="Regression Summary")
par(mfrow = c(2, 2))
plot(fit2)
```
